<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{jqr dsl introduction}
%\VignetteEncoding{UTF-8}
-->

```{r echo=FALSE}
knitr::opts_chunk$set(
	comment = "#>",
	collapse = TRUE,
	warning = FALSE,
	message = FALSE
)
```

Introduction to jqr DSL
=======================

There's low an high level interfaces in `jqr`. This vignette covers the high level interface, or DSL (domain specific language).

The `jqr` DSL uses a suite of functions to construct queries. Queries are constucted, then excuted internally with `jq()` after the last piped command.

You don't have to use pipes - they are optional. Though they do make sense in that you can build up queries easily, just as you would with `jq`, or any other tools, on the command line.

## The jqr DSL

* Execute jq
  * `jq` - execute jq
* Utility functions
  * `peek` - peek at query, without running it
  * `string` - give back character string
  * `combine` - combine pieces into proper JSON
* Identity functions
  * `dot` - takes its input and produces it unchanged as output.
  * `dotstr` - produces value at the key 'foo'
  * `index` - index to all elements, or elements by name or number
  * `indexif` - same as above, but shouldn't fail when not found
* Operations on keys, or by keys
  * `keys` - takes no input, and retrieves keys
  * `haskey` - checks if a json string has a key or keys
  * `del` - deletes provided keys
* Maths
  * `do` - arbitary math operations
  * `lengthj` - length, `j` on end to avoid collision with `base::length`
  * `sqrtj` - square root, `j` on end to avoid collision with `base::length`
  * `floorj` - returns the floor of its numeric input
  * `minj` - minimum element of input
  * `maxj` - maximum element of input
  * `add` - adds strings or numbers together
  * `map` - for any filter X, run X for each element of input array
* Manipulation operations
  * `join` - join strings
  * `splitj` - ...
  * `ltrimstr` -
  * `rtrimstr` -
  * `startswith` -
  * `endswith` -
  * `indices` -
  * `tojson` -
  * `fromjson` -
  * `tostring` -
  * `tonumber` -
  * `contains` -
  * `uniquej` -
  * `group` -

## Load jqr

```{r}
library("jqr")
```

## xxx

Examples soon...
