<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{jqr introduction}
%\VignetteEncoding{UTF-8}
-->

```{r echo=FALSE}
knitr::opts_chunk$set(
	comment = "#>",
	collapse = TRUE,
	warning = FALSE,
	message = FALSE
)
```

Introduction to jqr
===================

There's low and high level (or DSL [domain specific language]) interfaces in `jqr`.

## NSE vs. SE

Many functions in `jqr` have NSE as well as SE versions, where the NSE version for sorting an array is `sortj()` whereas the SE version is `sortj_()`. Some functions only have one version, and behave under SE rules.

## jqr low level interface

The low level and high level interfaces are unified via the function `jq()`. You can access the low leve interface by using `jq()` directly, passing a JSON string as the first parameter, the program (query) as the second, and [the flags](https://stedolan.github.io/jq/manual/#Invokingjq) as the third (by default no flags are passed).

For example, a JSON string could be `'{"a": 7, "b": 4}'`, and the program could be `.`, resulting in

```
{
    "a": 7,
    "b": 4
}
```

The program passed is exactly the same as you'd pass on the command line. Because this is simpl a replication of the command line in R, there is a higher level interface, or DSL.

## jqr DSL

The `jqr` DSL uses a suite of functions to construct queries that are excuted internally with `jq()` after the last piped command. We use some logic provided by Stefan Bache to determine whether the a function call is the last in a series of pipes, and if so, we run `jq()` on the JSON string passed, and the program/query passed.

You don't have to use pipes - they are optional. Though they do make things easier in that you can build up queries easily, just as you would with `jq`, or any other tools, on the command line.

* Execute jq
  * `jq` - execute jq
* Utility functions
  * `peek` - peek at query, without running it
  * `string` - give back character string
  * `combine` - combine pieces into proper JSON
* Identity functions
  * `dot` - takes its input and produces it unchanged as output.
  * `dotstr` - produces value at the key 'foo'
  * `index` - index to all elements, or elements by name or number
  * `indexif` - same as above, but shouldn't fail when not found
* Operations on keys, or by keys
  * `keys` - takes no input, and retrieves keys
  * `haskey` - checks if a json string has a key or keys
  * `del` - deletes provided keys
* Maths
  * `do` - arbitary math operations
  * `lengthj` - length, `j` on end to avoid collision with `base::length`
  * `sqrtj` - square root, `j` on end to avoid collision with `base::length`
  * `floorj` - returns the floor of its numeric input
  * `minj` - minimum element of input
  * `maxj` - maximum element of input
  * `add` - adds strings or numbers together
  * `map` - for any filter X, run X for each element of input array
* Manipulation operations
  * `join` - join strings on given separator
  * `splitj` - split string on separator argument
  * `ltrimstr` - removen given prefix string, if it starts with it
  * `rtrimstr` - removen given suffix string, if it starts with it
  * `startswith` - logical output, test if input start with foo
  * `endswith` - logical output, test if input ends with foo
  * `indices` - array with numeric indices where foo occurs in inputs
  * `tojson` - dump values to JSON
  * `fromjson` - parse JSON into values
  * `tostring` - convert to string
  * `tonumber` - convert to number
  * `contains` - logical output, determine if foo is in the input
  * `uniquej` - output unique set
  * `group` - groups the elements having the same .foo field into separate arrays
* Sort
  * `sortj` - sort an array
  * `reverse` - reverse an array
* Types
  * `type` - select elements by type
  * `types` - get type for each element
* Functions
  * `funs` - Define and use functions
* Variables
  * `vars` - Define variables to use later
* Recurse
  * `vars` - Search through a recursive structure - extract data from all levels
* Paths
  * `paths` - Outputs paths to all the elements in its input
* Range
  * `range` - Produce range of numbers
* Format strings
  * `at` - Format strings and escaping

## Load jqr

```{r}
library("jqr")
```

## Utility functions

Peek

```{r}
'{"a": 7}' %>% do(.a + 1) %>% peek
'[8,3,null,6]' %>% sortj %>% peek
```

String

```{r}
'{"a": 7}' %>% do(.a + 1) %>% string
'[8,3,null,6]' %>% sortj %>% string
```

Combine

```{r}
x <- '{"foo": 5, "bar": 7}' %>% select(a = .foo)
combine(x)
```

## Some examples

Index

```{r}
x <- '[{"message": "hello", "name": "jenn"}, {"message": "world", "name": "beth"}]'
x %>% index()
```

Sort

```{r}
'[8,3,null,6]' %>% sortj
```

reverse order

```{r}
'[1,2,3,4]' %>% reverse
```

Get multiple outputs for array w/ > 1 element

```{r eval=FALSE}
x <- '{"user":"stedolan","titles":["JQ Primer", "More JQ"]}'
x %>% index()
x %>% select(user, title = `.titles[]`)
x %>% select(user, title = `.titles[]`) %>% combine
```

join

```{r}
'["a","b,c,d","e"]' %>% join
'["a","b,c,d","e"]' %>% join(`;`)
```

endswith

```{r}
'["fo", "foo", "barfoo", "foobar", "barfoob"]' %>% index %>% endswith(foo)
```

contains

```{r}
'"foobar"' %>% contains("bar")
```

unique

```{r}
'[1,2,5,3,5,3,1,3]' %>% uniquej
```

Get type information for each element

```{r}
'[0, false, [], {}, null, "hello"]' %>% types
'[0, false, [], {}, null, "hello", true, [1,2,3]]' %>% types
```

Select elements by type

```{r}
'[0, false, [], {}, null, "hello"]' %>% index() %>% type(booleans)
```

Get keys

```{r}
str <- '{"foo": 5, "bar": 7}'
str %>% keys()
```

Delete by key name

```{r}
str %>% del(bar)
```

Check for key existence

```{r}
str3 <- '[[0,1], ["a","b","c"]]'
str3 %>% haskey(2)
str3 %>% haskey(1,2)
```

Select variables by name, and rename

```{r}
'{"foo": 5, "bar": 7}' %>% select(a = .foo)
```

More complicated `select()`, using the included dataset `githubcommits`

```{r}
githubcommits %>%
  index() %>%
  select(sha = .sha, name = .commit.committer.name)
```

Maths comparisons

```{r}
'[5,4,2,7]' %>% index() %>% do(. < 4)
'[5,4,2,7]' %>% index() %>% do(. > 4)
'[5,4,2,7]' %>% index() %>% do(. <= 4)
'[5,4,2,7]' %>% index() %>% do(. >= 4)
'[5,4,2,7]' %>% index() %>% do(. == 4)
'[5,4,2,7]' %>% index() %>% do(. != 4)
```

Sqrt

```{r}
'9' %>% sqrtj
```

Floor

```{r}
'3.14159' %>% floorj
```

Find minimum

```{r}
'[5,4,2,7]' %>% minj
'[{"foo":1, "bar":14}, {"foo":2, "bar":3}]' %>% minj
'[{"foo":1, "bar":14}, {"foo":2, "bar":3}]' %>% minj(foo)
'[{"foo":1, "bar":14}, {"foo":2, "bar":3}]' %>% minj(bar)
```

Find maximum

```{r}
'[5,4,2,7]' %>% maxj
'[{"foo":1, "bar":14}, {"foo":2, "bar":3}]' %>% maxj
'[{"foo":1, "bar":14}, {"foo":2, "bar":3}]' %>% maxj(foo)
'[{"foo":1, "bar":14}, {"foo":2, "bar":3}]' %>% maxj(bar)
```
